Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-11-21 papers/integration/omm_integ.tex
These changes add figures and text changes to (almost) complete the responses to the reviewers materials and results comments. Discussion re-write is still pending.

% Remove after review:
Note: Changes based on Reviewer 1's comments \rvw{are in this colour}. For
Reviewer 2 \rvwt{we use this colour}.

@@

% Parameterisation here?
\rvwt{The \ccg~model was parameterised by tuning the model to perform
a prosaccade task in which a fixed luminance point is fixated by teh
model. After a fixed duration, the fixation point was extinguished and
a target point of fixed luminance was presented. The model was tuned
so that the latency between the presentation of the target and the
initiation of an eye movement matched experimental data, while also
matching the electrophysiological evidence of activity in a variety of
brain regions. The tuning of the BG model attempted to preserve as
closely as possible the weights used in the original paper. Further
details on the parameterisation of the \ccg~model are
in \cite{cope_basal_2017}.}
% Differences between ccg and existing brain model or can I do that
% above? Better in the preceding text, I think and wherever it crops
% up.

@@

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in
Fig.~\ref{sbg}. \rvw{We use the word `channel' to mean a set of
populations of neurons which are involved in actuating a single
extraocular muscle. SBG channels are arranged in pairs, actuating
opposing muscles. There is one pair of channels which actuates the
superior and inferior rectus muscles, causing vertical rotations of
the eye in a roughly parasaggital plane (the eye moves up or
down). Another pair actuates the lateral and medial rectus muscles,
causing horizontal rotations of the eye. The third pair actuates the
superior and inferior oblique muscles which contribute to vertical as
well as oblique rotations.} Activity from the output layer of superior
colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

@@

The output signals of the brainstem are used to drive the
biomechanical model.  The latter is not only used to get tangible
feedback on the simulated saccades including motion trajectories, but
adds one more modelling dimension related to the inertial properties
of the eye plant including muscle properties.

@@

% Brain to SBG, seeing as that's how I explained it above.
A number of studies have considered the form of the connection between
the deeper layers of the superior colliculus and the saccadic burst
generator \citep{van_gisbergen_experimental_1985,ottes_visuomotor_1986,waitzman_superior_1991,groh_converting_2001,arai_two-dimensional_1994,goossens_dynamic_2006,tabareau_geometry_2007,van_opstal_linear_2008,goossens_optimal_2012},
which has become known as the spatial temporal transform (STT).  The
spatial aspect of the transform is thought to be implemented by a
weight-mapping \citep{tabareau_geometry_2007,arai_two-dimensional_1994}
and we follow this idea.
%
Arai and co-workers trained a 20x20 neural network model of the
superior colliculus to discover the weight map under the assumption of
2D Gaussian activation profiles~\citep{arai_two-dimensional_1994}.
%
The training approach of \cite{arai_two-dimensional_1994} was not
feasible in this study due to the length of time required to run our
model and its stochasticity, which meant multiple runs of the model
were necessary in order to generate output statistics.
%
\cite{tabareau_geometry_2007} wrote down a theoretical form of the weight
map, which follows from the mapping of \cite{ottes_visuomotor_1986}
and the assumption of invariant 2D Gaussian activity profiles in
SC. As they found it closely resembles the results
of \cite{arai_two-dimensional_1994}, and it is a simple formulation,
we considered it as the means to generate the six weight maps in our
own model.
%
One barrier to the use of the weight map
in \cite{tabareau_geometry_2007} was the \ccg~model's violation of the
\e{invariant integral hypothesis}. This states that the number of spikes
emitted by a neural element during a saccade (or in our model, the
integral of the neuron's output during the saccade) should be a
function only of its position within the hill of collicular
activity. That is, for any time-dependent hill of activity
$\mathcal{A}(\mathbf{z},t)$ at $\mathbf{z} = (r,\phi)$ on the collicular
surface, the integrated activity  $A_{\mathbf{x}}$ in an element at a
vector $\mathbf{x}$ away from $\mathbf{z}$ is
\begin{equation}
A_{\mathbf{x}} = \int_t \mathcal{A}(\mathbf{z}-\mathbf{x}, t)\;dt
\end{equation}
which is invariant for all $\mathbf{z}$. However, the very mapping on
which the \cite{tabareau_geometry_2007} result is based leads to a
very \e{variant} activity profile in the \ccg~model. A luminance of a
given size which excites activity near to the fovea causes activity in
a large number of neurons, whereas activity far from the fovea excites
a much smaller region. This effect is clearly demonstrated in
Fig.~\ref{fig:mapping} for equal sized targets both on and distal from
the fovea.
% Try to make this clearer for Reviewer 2, Methods Comments, point 5.

This led us to hypothesise that the retinotopic mapping be accompanied
by an associated widening projection field such that the hill of
activity in superior colliculus is invariant with position on the
collicular surface. There are a number of locations in the system in
which this widening projection field could exist. It could be
implemented in the projections between the retinal populations and the
superficial layer of SC along with the projection between the World
and the FEF population. However, this would affect activity within the
basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
presents evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{scdeep} (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor; within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

@@

To verify that our implementation of the brain model has the same
functionality as that reported in \cite{cope_basal_2017}, we
investigated the effect on saccadic response times of: target
eccentricity; and any gap or overlap between fixation off-time and
target on-time. We showed that the full model reproduces the `hockey
stick' shape shown in Fig. 7 of \cite{cope_basal_2017} for horizontal
[Fig.~\ref{lat_vs_all}(a)], vertical [Fig.~\ref{lat_vs_all}(b)] and
oblique saccades (not shown). The latency increases with eccentricity
far from the fovea because the retinotopic mapping reduces the
activity in the basal ganglia for more eccentric targets (this effect
is described in detail in \cite{cope_basal_2017}). Closer to the
fovea, the interaction between the foveal mask and the activity in FEF
again leads to reduced input into into the basal ganglia and an
increased time to achieve disinhibition in SNr.

@@

In another experiment, we probed the response of the model to double
step stimuli of the type described in \cite{becker_analysis_1979}. In
that work, the response of human subjects was investigated when shown
stimuli at 15\dg~and 30\dg~eccentricity with variable delay between
the stimuli. If the smaller eccentricity stimulus was shown first,
followed by the more distal on the same side of the field of view,
this was called a `staircase' presentation.
% If the more distal stimulus preceded the 15\dg~stimulus (again with
% both in the same hemifield), this was called `pulse undershoot'.
We carried out a `staircase' presentation, shown in
Fig.~\ref{doublesteps}, where our small eccentricity luminance was at
8\dg~and our more distal luminance was at 12\dg~(both to the right of
centre). We found that there was a critical time delay between the
luminances of about 30~ms. If they were presented with a delay smaller
than this value, then a single, slightly hypermetric saccade was
made. This response type is called a \e{final angle response}. A delay
greater than 30~ms between the stimuli would lead to double step
saccades (a so-called \e{initial angle response}), with the first
saccade arriving at 8\dg~(though with greater variability than
normal), and a second saccade being made to a location hypometric of
12\dg~after a pause of about 240~ms.  Fig.~\ref{doublesteps}(a) shows
the mean trajectories from 5 simulations of the staircase doublestep
presentation alongside the result for a single saccade to the final
angle of 12\dg. Dash-dot lines show $\pm$1 standard deviation about
the mean. The corresponding trajectories are shown in
Fig.~\ref{doublesteps}(b).

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./figures/ccg_model.png}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:ccg_model} Figure \arabic{figure}.}
{ \bol{a) The macroscopic architecture of the \ccg~model.} The main
nuclei modelled as brain systems are: basal ganglia (BG), frontal eye
fields (FEF), thalamus (THAL) and superior colliculus (SC). The
retinal input is presented via non-biologically plausible retinal
populations. The loops through basal ganglia, which define the
architecture, are shown with coloured lines: the cortical loop
(through FEF and THAL) in green and the sub-cortical loop (through SC
and THAL) in blue. Connections with arrowheads indicate excitatory
connections, those with circles are inhibitory. A red border indicates
that the box represents a sub-system of two or more populations; a
black box indicates (at least, within the context of the model) a
single neural population. The BG box is expanded in: \bol{b) The basal
ganglia model component.}
%
This shows a basal ganglia comprising striatum (Str\_D1 \& Str\_D2),
subthalamic nucleus (STN), globus pallidum externum (GPe) and
substantia nigra reticulata (SNr). The model has three action channels
shown as black boxes within each blue population border. Three
channels of cortical input to the BG are also depicted. Red indicates
the activation level of a given channel, helping to illustrate the
selection mechanism. For example, the channel indicated by the
leftmost bar has a high salience (cortical input) and excites activity
in Str\_D1 which then inhibits the leftmost bar in SNr. The diffuse
projection from STN is equivalent to summing its projections
channel-wise, and then projecting the sum to all channels of its
target populations (the blue arrows indicate that all channels of GPe
and SNr are targeted by the connection). Dopaminergic modulation of
the inputs to the striatum are indicated by the blue circles labelled
`d' and the dotted lines. The SNr sends inhibitory output projections
to its targets.}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.5\textwidth]{./figures/eye_screen_plus.png}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:bioeye} Figure \arabic{figure}.}
{ The biomechanical eye. a) an OpenSim rendering of a pair of
biomechanical eyes showing the positions of the extraocular
muscles. b) OpenSim rendering of biomechanical eye within a head
model c) Top-down schematic cross-sectional view of the biomechanical eye situated
within a spherical screen, with a horizontal rotation towards a
luminances at an angle of $-\theta_{y}$ about the $y$ axis, which
points up out of the page.}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.5\textwidth]{./figures/SC_to_brainstem.png}
\end{center}
\textbf{\refstepcounter{figure}\label{scdeep} Figure \arabic{figure}.}
{ Showing the additional deep layer of superior colliculus (SC\_deep2)
and the output layer (SC\_avg, named for the fact that in an earlier
version of the model, it received the output of the centroid of
SC\_deep).  The widening Gaussian projection is shown as the arrow
between SC\_deep and SC\_deep2.}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./figures/lat_vs_everything.png}
\end{center}
\textbf{\refstepcounter{figure}\label{lat_vs_all} Figure \arabic{figure}.}
{ Exploring saccade latencies. a) Latency to first movement as a
function of target eccentricity for horizontal targets. b) Latency
vs. eccentricity for vertical targets. c) Latency vs. gap at three
different luminance values. d) The effect of the dopamine parameter on
saccade latencies in gap, step and overlap conditions, for two
different target luminances.  e) Saccade vs. luminance showing gradual
transition between reflexive and express behaviour.
%f) Saccade duration vs. saccade eccentricity for horizontal
%(rightwards), vertical (downwards) and 45\dg~oblique (down \& right) saccades.
}
\end{figure}

>>>>

% Remove after review:
\bol{Note:} Changes based on Reviewer 1's comments \rvw{are in this colour}. For
Reviewer 2 \rvwt{we use this colour}.

@@

% Parameterisation here?
\rvwt{The \ccg~model was parameterised by tuning the model to perform
a prosaccade task in which a fixed luminance point is fixated by teh
model. After a fixed duration, the fixation point was extinguished and
a target point of fixed luminance was presented. The model was tuned
so that the latency between the presentation of the target and the
initiation of an eye movement matched experimental
data \citep{reulen_latency_1984}, while also matching the
electrophysiological evidence of activity in a variety of brain
regions. The tuning of the BG model attempted to preserve as closely
as possible the weights used in the original paper. Further details on
the parameterisation of the \ccg~model are in \cite{cope_basal_2017}.}
% Differences between ccg and existing brain model or can I do that
% above? Better in the preceding text, I think and wherever it crops
% up.

@@

% Was looking for somewhere for this, possibly:
%The precision of the model to distinguish between
%targets, especially towards the periphery of the field of view was
%limited by the 50$\times$50 size of the neural layers. In order to optimise
%this precision, both the \ccg~model and our implementation of
%it had their effective eccentricity range limited. Our implementation
%was limited more than the Cope-Chambers model, because hills of
%activity in SC_deep had to be kept fully within each layer. When a
%hill `falls off the edge' of the layer, the amount of activation being
%passed to the SBG is reduced and the saccades become inaccurate. I
%didn't want to investigate effects at the edge of the SC, so I had an
%effective limit of the model of about 22 degrees of eccentricity for a
%saccade.

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in
Fig.~\ref{sbg}. \rvw{In the brainstem model, we use the word `channel'
to mean a set of populations of neurons which are involved in
actuating a single extraocular muscle. SBG channels are arranged in
pairs, actuating opposing muscles. There is one pair of channels which
actuates the superior and inferior rectus muscles, causing vertical
rotations of the eye in a roughly parasaggital plane (the eye moves up
or down). Another pair actuates the lateral and medial rectus muscles,
causing horizontal rotations of the eye. The third pair actuates the
superior and inferior oblique muscles which contribute to vertical as
well as oblique rotations.} Activity from the output layer of superior
colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

@@

The output signals of the brainstem's \rvwt{motoneuron (MN)
populations} are used to drive the biomechanical model. \rvwt{The MN
output signal in each brainstem channel is normalised in the range [0
1] and represents the mean firing rate of the neurons that innervate
the extraocular muscle for that channel.} The biomechanics are used
not only to get tangible feedback on the simulated saccades including
motion trajectories, but to add one more modelling dimension related
to the inertial properties of the eye plant including muscle
properties.

@@

% Brain to SBG, seeing as that's how I explained it above.
A number of studies have considered the form of the connection between
the deeper layers of the superior colliculus and the saccadic burst
generator \citep{van_gisbergen_experimental_1985,ottes_visuomotor_1986,waitzman_superior_1991,groh_converting_2001,arai_two-dimensional_1994,goossens_dynamic_2006,tabareau_geometry_2007,van_opstal_linear_2008,goossens_optimal_2012},
which has become known as the spatial temporal transform (STT).  The
spatial aspect of the transform is thought to be implemented by a
weight-mapping \citep{tabareau_geometry_2007,arai_two-dimensional_1994}
and we follow this idea.
%
Arai and co-workers trained a 20x20 neural network model of the
superior colliculus to discover the weight map under the assumption of
2D Gaussian activation
profiles~\citep{arai_two-dimensional_1994}---\rvwt{that is, they
assumed that the activity in superior colliculus for any saccade was a
size-invariant 2D Gaussian hill of activity}.
%
The training approach of \cite{arai_two-dimensional_1994} was not
feasible in this study due to the length of time required to run our
model and its stochasticity, which meant multiple runs of the model
were necessary in order to generate output statistics.
%
\cite{tabareau_geometry_2007} wrote down a theoretical form of the weight
map, \rvwt{obtained by inverting the mapping} of \cite{ottes_visuomotor_1986}
and the assumption of invariant 2D Gaussian activity profiles in SC,
\rvwt{which is equivalent to:}
%
\begin{equation} \label{eq:tabareau_eq3}
w(r,\phi) = i\;e\,^{jr}\;\sin\left(l\phi + k\right)
\end{equation}
\rvwt{where $r$ and $\phi$ are co-ordinates on the collicular map and
$i$, $j$, $k$ and $l$ are parameters of the
function (compare with Eq.~3 of \cite{tabareau_geometry_2007}).}
%
As they found it closely resembles the results
of \cite{arai_two-dimensional_1994}, and it is a simple formulation,
we considered it as the means to generate the six weight maps in our
own model.
%
One barrier to the use of this weight map was the \ccg~model's
violation of the
\e{invariant integral hypothesis}. This states that the number of spikes
emitted by a neural element during a saccade (or in our model, the
integral of the neuron's output during the saccade) should be a
function only of its position within the hill of collicular
activity. That is, for any time-dependent hill of activity
$\mathcal{A}(\mathbf{z},t)$ at $\mathbf{z} = (r,\phi)$ on the collicular
surface, the integrated activity  $A_{\mathbf{x}}$ in an element at a
vector $\mathbf{x}$ away from $\mathbf{z}$ is
\begin{equation}
A_{\mathbf{x}} = \int_t \mathcal{A}(\mathbf{z}-\mathbf{x}, t)\;dt
\end{equation}
which is invariant for all $\mathbf{z}$. \rvwt{This requirement is
fulfilled by spatially invariant 2D Gaussian profiles, whose
time-course (how quickly they grow and then diminish) is always the
same.}

However, the very mapping on which the \cite{tabareau_geometry_2007}
result is based leads to a very \e{variant} activity profile \rvwt{in
the SC\_deep layer} of the \ccg~model. A luminance of a given size
which excites activity near to the fovea causes activity in a large
number of neurons in each retinotopic layer, whereas activity far from
the fovea excites a much smaller region. This effect is clearly
demonstrated in Fig.~\ref{fig:mapping} for equal sized targets both on
and distal from the fovea.

% Try to make this clearer for Reviewer 2, Methods Comments, point 5.
\rvwt{To understand the need for this invariance,
consider the effect of a 2D Gaussian hill in SC\_deep which elicits a
successful horizontal saccade of 10\dg. Activity from the 2D Gaussian,
passing through the weight maps will excite the superior and inferior
rectus channels by an equal, balanced amount, so these cancel out,
allowing the eye movement to be horizontal. The amount of activation
passed to the lateral rectus muscle results from a convolution of the
Gaussian and the exponential component of the weight map relationship
in Eq.~\ref{eq:tabareau_eq3}. If the Gaussian hill now appears further
along the collicular surface, coding for a 20\dg~saccade, \e{and also
becomes smaller}, we can still argue that the vertical component
signals to superior/inferior rectus muscles will cancel out, and we
could imagine that the exponential component of
Eq.~\ref{eq:tabareau_eq3} is correctly parameterised to compensate for
the smaller hill. However. Now consider a 2-D Gaussian hill which
codes for a 10\dg saccade which is `up, and to the right' in equal
proportions. That means that the hill will sit on the boundary between
the weight maps for the `up', and the `right' muscles. Now, if the
hill moves to the $r=$20\dg~location on the colliculus, \e{and also
reduces in size}, it will excite only the periphery of the sine; the
exponential increase of the map along $r$ is not guaranteed to
compensate for the reduction in the convolution of the Gaussian hill
and the sinusoidal component of the weight map.}
% Add figure for this.

This led us to hypothesise that the retinotopic mapping \rvwt{to the
SBG be preceded} by an associated widening projection field such that
the hill of activity in \rvwt{a `final' deep layer of superior
colliculus is invariant with position on the map}. There are a number of
locations in the system in which this widening projection field could
exist. It could be implemented in the projections between the retinal
populations and the superficial layer of SC along with the projection
between the World and the FEF population. However, this would affect
activity within the basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
presents evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{scdeep}(b) (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor; within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

@@

To verify that our implementation of the brain model has the same
functionality as that reported in \cite{cope_basal_2017}, we
investigated the effect on saccadic response times of: target
eccentricity; and any gap or overlap between fixation off-time and
target on-time. We showed that the full model reproduces the `hockey
stick' shape shown in Fig. 7 of \cite{cope_basal_2017} for horizontal
[Fig.~\ref{lat_vs_all}(a)], vertical [Fig.~\ref{lat_vs_all}(b)] and
oblique saccades (not shown). The latency increases with eccentricity
far from the fovea because the retinotopic mapping reduces the
activity in the basal ganglia for more eccentric targets (see
Fig.~\ref{mapping}). Closer to the fovea, the \rvwt{effect of} the
foveal mask \rvwt{on} the activity in FEF again leads to reduced input
into the basal ganglia and an increased time to achieve disinhibition
in SNr.

@@

In another experiment, we probed the response of the model to double
step stimuli of the type described in \cite{becker_analysis_1979}. In
that work, the response of human subjects was investigated when shown
stimuli at 15\dg~and 30\dg~eccentricity with variable delay between
the stimuli. If the smaller eccentricity stimulus was shown first,
followed by the more distal on the same side of the field of view,
this was called a `staircase' presentation.
% If the more distal stimulus preceded the 15\dg~stimulus (again with
% both in the same hemifield), this was called `pulse undershoot'.
We carried out a `staircase' presentation, shown in
Fig.~\ref{doublesteps}, where our small eccentricity luminance was at
8\dg~and our more distal luminance was at 12\dg~(both to the right of
centre). \rvwt{The stimuli could not be presented at 15\dg and 30\dg
to match the experiment, because 30\dg saccades were outside the range
of the model.}

We found that there was a critical time delay between the
luminances of about 30~ms. If they were presented with a delay smaller
than this value, then a single, slightly hypermetric saccade was
made. This response type is called a \e{final angle response}. A delay
greater than 30~ms between the stimuli would lead to double step
saccades (a so-called \e{initial angle response}), with the first
saccade arriving at 8\dg~(though with greater variability than
normal), and a second saccade being made to a location hypometric of
12\dg~after a pause of about 240~ms.  Fig.~\ref{doublesteps}(a) shows
the mean trajectories from 5 simulations of the staircase doublestep
presentation alongside the result for a single saccade to the final
angle of 12\dg. Dash-dot lines show $\pm$1 standard deviation about
the mean. The corresponding trajectories are shown in
Fig.~\ref{doublesteps}(b).

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=\textwidth]{./figures/ccg_model.png}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:ccg_model} Figure \arabic{figure}.}
{ \bol{a) The macroscopic architecture of the \ccg~model.} The main
nuclei modelled as brain systems are: basal ganglia (BG), frontal eye
fields (FEF), thalamus (THAL) and superior colliculus (SC). The
retinal input is presented via non-biologically plausible retinal
populations. The loops through basal ganglia, which define the
architecture, are shown with coloured lines: the cortical loop
(through FEF and THAL) in green and the sub-cortical loop (through SC
and THAL) in blue. Connections with arrowheads indicate excitatory
connections, those with circles are inhibitory. A red border indicates
that the box represents a sub-system of two or more populations; a
black box indicates (at least, within the context of the model) a
single neural population. The BG box is expanded in: \bol{b) The basal
ganglia model component.}
%
This shows a basal ganglia comprising striatum (Str\_D1 \& Str\_D2),
subthalamic nucleus (STN), globus pallidum externum (GPe) and
substantia nigra reticulata (SNr). The model has three action channels
shown as black boxes within each blue population border. Three
channels of cortical input to the BG are also depicted. Red indicates
the activation level of a given channel, helping to illustrate the
selection mechanism. For example, the channel indicated by the
leftmost bar has a high salience (cortical input) and excites activity
in Str\_D1 which then inhibits the leftmost bar in SNr. The diffuse
projection from STN is equivalent to summing its projections
channel-wise, and then projecting the sum to all channels of its
target populations (the blue arrows indicate that all channels of GPe
and SNr are targeted by the connection). Dopaminergic modulation of
the inputs to the striatum are indicated by the blue circles labelled
`d' and the dotted lines. The SNr sends inhibitory output projections
to its targets. \bol{c) 2D Gaussian weights.} The `GaussianKernel'
connectivity pattern is based on the horizontal distance between the
location of a neural element in one layer, and the location of a
target neural element in the target layer for the connection. The
potential weight of the connection is given by a 2D Gaussian function,
which is maximum for the target neuron which exactly corresponds to
the source neuron, and drops down for target neurons which are
horizontally displaced from the source neuron. A threshold is applied
to avoid a computationally expensive all-to-all connectivity (with
most of these connections having negligible weight values).  If the
weight is non-zero, then a connection is made from source to target,
otherwise no connection is made. \bol{d) Gaussian connectivity.} This
image shows connectivity (green rays) from two source neurons (in
Str\_D2, brown spheres) to target neurons (in GPe, blue spheres). The
circular connectivity pattern is seen. This does not show the weights,
which reduce `towards the edge of the circle' and follow the
relationship shown in (c).}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.5\textwidth]{./figures/eye_screen_plus.png}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:bioeye} Figure \arabic{figure}.}
{ The biomechanical eye. a) an OpenSim rendering of a pair of
biomechanical eyes showing the positions of the extraocular
muscles. Note that i) volume visualization of muscles should not be
confused with FEM muscle models; it is provided for user feedback
purposes, i.e. shape and color change depending on the muscle
activation, ii) superior and inferior oblique are visualized up to
their respective muscle pulleys. b) OpenSim rendering of biomechanical
eye within a head model c) Top-down schematic cross-sectional view of
the biomechanical eye situated within a spherical screen, with a
horizontal rotation towards a luminances at an angle of $-\theta_{y}$
about the $y$ axis. The $y$ axis points up, out of the page.}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.9\textwidth]{./figures/scdeep_plus.png}
\end{center}
\textbf{\refstepcounter{figure}\label{scdeep} Figure \arabic{figure}.}
{ a) \& b) Tabareau-style weight maps for `left' and `down' components
of a saccade. c) The two weight maps in a) and b) shown on the same
graph, viewed from above. Circles show the locations of potential
hills of activity. Purple, dashed circles encode saccades down;
red circles encode saccades with both a left and a down component. d)
Showing the additional deep layer of superior colliculus (SC\_deep2)
and the output layer (SC\_avg, named for the fact that in an earlier
version of the model, it received the output of the centroid of
SC\_deep).  The widening Gaussian projection is shown as the arrow
between SC\_deep and SC\_deep2.}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.8\textwidth]{./figures/lat_vs_everything.png}
\end{center}
\textbf{\refstepcounter{figure}\label{lat_vs_all} Figure \arabic{figure}.}
{ Exploring saccade latencies. a) Latency to first movement as a
function of target eccentricity for horizontal targets. b) Latency
vs. eccentricity for vertical targets. c) Latency vs. gap at three
different luminance values. The data are shown alongside
the \ccg~model results from Fig.~5 of \cite{cope_basal_2017} in blue,
and the experimental results used in that model in red. The fixation
luminance for the \ccg~curve was 0.5, the target lumianance 0.6 and
the target eccentricity was 8\dg. The difference between
the \ccg~model data and the data from the current model results from
the different mechanism by which activity in SC\_deep causes a
movement, the differing target angle and the reduced fixation
luminance used here. d) The effect
of the dopamine parameter on saccade latencies in gap, step and
overlap conditions, for two different target luminances.  e) Saccade
vs. luminance showing gradual transition between reflexive and express
behaviour.
%f) Saccade duration vs. saccade eccentricity for horizontal
%(rightwards), vertical (downwards) and 45\dg~oblique (down \& right) saccades.
}
\end{figure}
